Wait-free Queues with Polylogarithmic Step Complexity

### 2022-09-21 thesis/thesis.tex
minor fix

A block is an object storing some statistics, as described in Algorithm Queue. A block in a node's blocklist implicitly represents a set of operations.
\begin{definition}[Ordering of a block in a node]
Let $b$ be \nf{$n$.blocks[$i$]} and $b^\prime$ be \nf{$n$.blocks[$j$]}. We call $i$ the \emph{index} of block $b$. Block $b$ is \it{before} block $b^\prime$ in node \nf{n} if and only if $i<j$. We define \it{the prefix for} $b$ to be the blocks in \tt{n.blocks[0..$i$]}.
\end{definition}

@@

Now we define the ordering of operatiobs stored in each node. In the non-root nodes we only need ordering of operations of a type among themselves. Processes are numbered from 1 to $p$ and leaves of the tree are assigned from left to right. We will show in Lemma \ref{blockSize} that there is at most one operation from each process in a given block.
\begin{definition} [Ordering of operations inside the nodes] \label{ordering}

\begin{itemize}
  \item $E(n,b)$ is the sequence of enqueue operations in $ops(n\nf{.blocks[}b\nf{]})$ defined recursively as follows. $E(leaf,b)$ is the single enqueue operation in $ops(leaf\nf{.blocks[}b\nf{]})$ or an empty sequence if $leaf\nf{.blocks[}b\nf{].num\sub{enq}}=0$. If $n$ is an internal node, then
\begin{align*} 
E(n,b) =&  E(n\nf{.left},n\nf{.blocks[}b-1\nf{].end\sub{left}}+1)\cdot E(n\nf{.left},n\nf{.blocks[}b-1\nf{].end\sub{left}}+2)\cdots E(n\nf{.left},n\nf{.blocks[}b\nf{].end\sub{left}})\cdot \\ 
&E(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+1)\cdot E(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+2)\cdots E(n\nf{.right},n\nf{.blocks[}b\nf{].end\sub{right}})
\end{align*}
  \item $E_i(n,b)$ is the $i$th enqueue in $E(n,b)$.
\item The order of the enqueue operations in the node $n$ is $E(n)=E(n,1)\cdot E(n,2)\cdot E(n,3)\cdots$
\item $E_i(n)$ is the $i$th enqueue in $E(n)$.
  \item $D(n,b)$ is the sequence of dequeue operations in $ops(n\nf{.blocks[}b\nf{]})$ defined recursively as follows. $D(leaf,b)$ is the single dequeue operation in $ops(leaf\nf{.blocks[}b\nf{]})$ or an empty sequence if $leaf\nf{.blocks[}b\nf{].num\sub{deq}}=~0$. If $n$ is an internal node, then
\begin{align*} 
D(n,b) =&  D(n\nf{.left],n\nf{.blocks[}b-1\nf{].end\sub{left}}+1)\cdot D(n\nf{.left}, n\nf{.blocks[}b-1\nf{].end\sub{left}}+2)\cdots D(n\nf{.left}, n\nf{.blocks[}b\nf{].end\sub{left}})\cdot \\ 
&D(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+1)\cdot D(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+2)\cdots D(n\nf{.right},n\nf{.blocks[}b\nf{].end\sub{right}})
\end{align*}
    \item $D_i(n,b)$ is the $i$th enqueue in $D(n,b)$.
\item The order of the dequeue operations in the node $n$: $D(n)=D(n,1)\cdot D(n,2)\cdot D(n,3)...$
\item $D_i(n)$ is the $i$th dequeue in $D(n)$.
\end{itemize}
\end{definition}

@@

%\begin{lemma}[Search Ranges]\label{search}
%  Preconditions of all invocation of \nf{BSearch} are satisfied.
%\end{lemma}
%\begin{proof}
%  
%Line 83: \nf{Get(i)} is called if the result of a dequeue is not null. The search is among all blocks in the root.
%
%Line 88: This search tries to find the ith enqueue, knowing that it is in the left child. Search is done over the left subblocks. The start and end of the range are followed by definition. Line 92 is the same.
%
%Line 101: Here, the goal is to find the superblock. We know the distance between answer and the \nf{super[i]} is at most $p$, since at most $p$ processes could die.
%
%\end{proof}
\begin{definition}
   Assume the operations in $L$ are applied on an empty queue. If element of \nf{enqueue e} is the response to  \nf{dequeue d} then we say \nf{R(d)=e}. If \nf{d} 's response id \nf{null} (queue is empty) then \nf{R(d)=null}.
\end{definition}

>>>>

A block is an object storing some statistics, as described in Algorithm Queue. A block in a node implicitly represents a set of operations.
\begin{definition}[Ordering of a block in a node]
Let $b$ be \nf{$n$.blocks[$i$]} and $b^\prime$ be \nf{$n$.blocks[$j$]}. We call $i$ the \emph{index} of block $b$. Block $b$ is \it{before} block $b^\prime$ in node \nf{n} if and only if $i<j$. We define \it{the prefix for} $b$ to be the blocks in \tt{n.blocks[0..$i$]}.
\end{definition}

@@

Now we define the ordering of operatiobs stored in each node. In the non-root nodes we only need ordering of operations of a type among themselves. Processes are numbered from 1 to $p$ and leaves of the tree are assigned from left to right. We will show in Lemma \ref{blockSize} that there is at most one operation from each process in a given block.
\begin{definition} [Ordering of operations inside the nodes] \label{ordering}
\end{definition}
\begin{itemize}
  \item $E(n,b)$ is the sequence of enqueue operations in $ops(n\nf{.blocks[}b\nf{]})$ defined recursively as follows. $E(leaf,b)$ is the single enqueue operation in $ops(leaf\nf{.blocks[}b\nf{]})$ or an empty sequence if $leaf\nf{.blocks[}b\nf{].num\sub{enq}}=0$. If $n$ is an internal node, then
\begin{align*} 
E(n,b) =&  E(n\nf{.left},n\nf{.blocks[}b-1\nf{].end\sub{left}}+1)\cdot E(n\nf{.left},n\nf{.blocks[}b-1\nf{].end\sub{left}}+2)\cdots E(n\nf{.left},n\nf{.blocks[}b\nf{].end\sub{left}})\cdot \\ 
&E(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+1)\cdot E(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+2)\cdots E(n\nf{.right},n\nf{.blocks[}b\nf{].end\sub{right}})
\end{align*}
  \item $E_i(n,b)$ is the $i$th enqueue in $E(n,b)$.
\item The order of the enqueue operations in the node $n$ is $E(n)=E(n,1)\cdot E(n,2)\cdot E(n,3)\cdots$
\item $E_i(n)$ is the $i$th enqueue in $E(n)$.
  \item $D(n,b)$ is the sequence of dequeue operations in $ops(n\nf{.blocks[}b\nf{]})$ defined recursively as follows. $D(leaf,b)$ is the single dequeue operation in $ops(leaf\nf{.blocks[}b\nf{]})$ or an empty sequence if $leaf\nf{.blocks[}b\nf{].num\sub{deq}}=~0$. If $n$ is an internal node, then
\begin{align*} 
D(n,b) =&  D(n\nf{.left},n\nf{.blocks[}b-1\nf{].end\sub{left}}+1)\cdot D(n\nf{.left}, n\nf{.blocks[}b-1\nf{].end\sub{left}}+2)\cdots D(n\nf{.left}, n\nf{.blocks[}b\nf{].end\sub{left}})\cdot \\ 
&D(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+1)\cdot D(n\nf{.right},n\nf{.blocks[}b-1\nf{].end\sub{right}}+2)\cdots D(n\nf{.right},n\nf{.blocks[}b\nf{].end\sub{right}})
\end{align*}
    \item $D_i(n,b)$ is the $i$th enqueue in $D(n,b)$.
\item The order of the dequeue operations in the node $n$: $D(n)=D(n,1)\cdot D(n,2)\cdot D(n,3)...$
\item $D_i(n)$ is the $i$th dequeue in $D(n)$.
\end{itemize}

@@

\begin{definition}
   Assume the operations in $L$ are applied on an empty queue. If element of \nf{enqueue e} is the response to  \nf{dequeue d} then we say \nf{R(d)=e}. If \nf{d} 's response id \nf{null} (queue is empty) then \nf{R(d)=null}.
\end{definition}
